---
title: "scfmRegime"
author: "Module Author"
date: "03 August 2015"
output: pdf_document
---

Module documentation should be written so that others can use your module.
This is a template for module documentation, and should be changed.

Parameter _empiricalMaxSizeFactor_ is used to derive max from mean fire size when the statistical estimator fails; 

Hi Steve,

I did a bit more digging. I mixed the numbers up. The 1.3 value was calculated from the hexagons (10,000km^2, n=484), and the 1.2 value from the ecoregions (n=62). 
## Plotting

Write what is plotted.

## Saving

Write what is saved.

## Input data

## Output data

## Anticipated linkages to other modules

## Other Markdown help
For help writing in RMarkdown, see http://rmarkdown.rstudio.com/. We have also included
The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

output:
rmarkdown::html_vignette:
css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side.

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

output:
rmarkdown::html_vignette:
fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))

