---
title       : "Exercises: `SpaDES` in action"
author      : "Alex M Chubaty & Eliot McIntire"
date        : "December 7, 2016"
output: 
  html_document: 
    number_sections: yes
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, eval = FALSE)
```
# `SpaDES` sample modules

1.  Which modules are included with the `SpaDES` package?

    ```{r sampleModuleDir}
    ## Locate the `SpaDES` sample modules
    sampleModuleDir <- system.file('sampleModules', package = 'SpaDES')
    
    ## identify the modules in this directory
    dir(sampleModuleDir)
    ```

2. What do each of the sample modules do? *I.e.*, what does each one demonstrate?

    ```{r sampleModuleFiles}
    ## open the sample module files to learn more about each module
    sampleModuleFiles <- list.files(sampleModuleDir, recursive = TRUE, full.names = TRUE)
    lapply(sampleModuleFiles, file.edit)
    ```

# `LCC2005` vegetation dynamics modules

Download and examine the `LCC2005` vegetation dynamics modules to learn more about them.

```{r LCC2005ModuleFiles, eval=FALSE}
## specify where to save files
moduleDir <- file.path(tempdir(), 'SpaDES_modules')

## download all necessary modules
downloadModule('LCC2005', path = moduleDir, data = TRUE)

## look at module or helper .Rmd file
openModules('LCC2005', path = moduleDir) 
```

# `wolfAlps` module

Download and examine the `wolfAlps` individual based model to learn more about it.

```{r wolf}
downloadModule("wolfAlps", path = moduleDir, data = TRUE)
openModules("wolfAlps", path = moduleDir)
```

# `SpaDES` and `shiny`

If time permits, feel free to play around with the proof-of-concept shiny apps:

- https://spades.shinyapps.io/ForestChange_ProofOfConcept/
- http://w-vic-a105388.pfc.forestry.ca/SpaDESUI-forEliot-Oct2016/ (only available inside CFS firewall)
