---
title       : "Dynamic modelling in SpaDES - a simple example"
author      : "Ceres Barros"
date        : "January 26, 2020"
output: 
  html_document: 
    number_sections: yes
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, eval = FALSE)
```

# Dynamic modelling in SpaDES - a simple example

## A Lotka-Volterra-type model

Lotka-Volterra predator-prey models are simple models that simulate the population dynamics of a prey and a predator, relying on quite a few assumptions. In this case we assume that:
* in the absence of predators, the prey population density follows a logistic growth equation, that depends on its intrinsic growth rate, *r*, and the environment's carrying capacity *K*.
* in the presence of predators, the number of prey consumed is a function of the hunting/consumption efficiency if the predator, *beta*, and the number of predators.
* the predator population density is solely dependent on the availability of prey, the predator's ability to convert prey into energy (predation efficiency), *delta*, and the intrinsic death rate of the predators in the absence of prey, *gamma*.

Th code bellow simulates these dynamics in a *for-loop* in the following steps:

1. Load necessary libraries;
2. Define parameters entering the model;
3. Loop the model a number of times and save the results in two vectors;
4. Plot the results.

You'll notice that the model is extremely sensitive to parameter values - this is a well-know feature of the Lotka-Volterra model. The ones bellow will ensure an oscillating equilibrium between prey and predator populations.

```{r lotka-volterradynamics things}
## load libraries
library(ggplot2)
library(data.table)

## define parameter values
r <- 0.1
K <- 100
beta <- 0.02
delta <- 0.5
gamma <- 0.1
time <- 100

## make 2 storage vectors for prey and predator densities
prey <- pred <- rep(NA, time)  

## set the inital population values
prey[1] <- 10
pred[1] <- 2

## The model in a loop - note that we need to start from the second time step
for(i in 2:time) {
  ## prey growth
  prey[i] <- prey[i-1]*exp(r*(1 - prey[i-1]/K))
  if (prey[i] < 0) prey[i] <- 0
            
  ## predation
  preyConsumed <- beta*prey[i-1]*pred[i-1]
  prey[i] <- prey[i] - preyConsumed
  
  ## predator growth    
  pred[i] <- pred[i-1] + delta*preyConsumed - gamma*pred[i-1]
  if (pred[i] < 0) pred[i] <- 0
}

## plot results:
plotData <- data.table(prey = prey, pred = pred, time = 1:time)
plotPrey <- ggplot(plotData) +
  geom_line(aes(x = time, y = prey), col = "blue", size = 1)
plotPred <- ggplot(plotData) +
  geom_line(aes(x = time, y = pred), col = "red", size = 1)

Plot(plotPrey, plotPred)
```

## Convert this to a SpaDES module

Now to the *actual* exercise. Try to convert this model into a 2-module SpaDES project/model. 
Here are a few tips:

1. Use `newModule()` (twice) to create the blank templates for your predator and prey population modules.
2. Open the `.R` files that show up in the module folders.
3. Fill in metadata:

  - Indicate libraries that are needed
  - Indicate parameters and provide their defaults
  - Indicate objects that are expected as an inputs, and objects that 
  are created as outputs

4. Fill in `.inputObjects` section to create the needed inputs.
5. Put all the code that creates the `distFromSource` and `establish` in the "init" event.
6. Put Plot stuff in `plot` event.

## Build a controller for this module.

7. Open the `.Rmd` file that we will use.
8. Run one line at a time, confirming that it will run your model. You will likely need to update the module path so it knows where to find the module on your computer.

# Advanced -- Add space
Try to make these modules spatially explicit if you have extra time. Hints:
1. You can use raster layers, instead of vectors, to save/track `preyPop` and `predPop` and use raster-based operations to calculate how many prey and consumed in each pixel.
2. You can use `data.tables` to speed up the computations, where each line corresponds to a pixel in a template raster and the columns are `preyPop`, `predPop` and `consumedPrey`. Remaining paratemers can be in a separate `data.table`, or simply left as numeric vectors. Then, for plotting, you can create and replace values in rasters to show spatial variation in predator and prey densities.

# One possible solution:
[For prey module](https://github.com/PredictiveEcology/SpaDES-modules/blob/master/modules/prey/prey.R)
[For predator module](https://github.com/PredictiveEcology/SpaDES-modules/blob/master/modules/predator/predator.R)

This solution follows the a-spatial option.
