---
title       : "Caching and debugging"
author      : "Ceres Barros"
date        : "January 2020"
output: 
  html_document: 
    number_sections: yes
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, eval = TRUE)
```

# Some basic Caching lessons

Caching is the ability to save some sort of output from an operation, and then retrieve these outputs when the operation is repeated in the *same way* - meaning the inputs of this operation and the actual tasks it performs are unchanged.

Caching becomes fundamental when we can expect to re-run operations several times, particularly if they they a while to compute each time. Some examples of these operations are:
- downloading data
- (spatial) data processing/munging
- fitting statistical models to large datasets, or that are complex in nature
- running simulations with no stochasticity

`SpaDES` (via the `reproducible` package) offers a number of functions that make caching these operations a lot easier for non-programmers. Two fundamental ones are `Cache` and `prepInputs`.

```{r LCC2005_1, eval = FALSE}
library(reproducible)
library(sp)
library(raster)

# Create a "study area"
coords <- structure(c(-122.98, -116.1, -99.2, -106, -122.98, 59.9, 65.73, 63.58, 54.79, 59.9),
                    .Dim = c(5L, 2L))
Sr1 <- Polygon(coords)
Srs1 <- Polygons(list(Sr1), "s1")
StudyArea <- SpatialPolygons(list(Srs1), 1L)
crs(StudyArea) <- "+init=epsg:4326 +proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

## from ?prepInputs
dPath <- file.path(tempdir(), "LCC")
lcc2005Filename <- file.path(dPath, "LCC2005_V1_4a.tif")
url <- file.path("ftp://ftp.ccrs.nrcan.gc.ca/ad/NLCCLandCover",
                 "LandcoverCanada2005_250m/LandCoverOfCanada2005_V1_4.zip")
LCC2005 <- prepInputs(url = url,
                      destinationPath = asPath(dPath),
                      useSAcrs = TRUE,  ## a reproducible::postProcess argument
                      studyArea = StudyArea)
```

Now do it again. Notice any difference?
```{r LCC2005_2, eval = FALSE}
LCC2005 <- prepInputs(url = url,
                      destinationPath = asPath(dPath),
                      useSAcrs = TRUE,  ## a reproducible::postProcess argument
                      overwrite = TRUE, ## a reproducible::postProcess argument
                      studyArea = StudyArea)
```

Now try wrapping the previous operation in `Cache` call, and run it twice. Notice differences in speed.
```{r LCC2005_3, eval = FALSE}
LCC2005 <- Cache(prepInputs,
                 url = url,
                 destinationPath = asPath(dPath),
                 useSAcrs = TRUE,
                 overwrite = TRUE,
                 studyArea = StudyArea)
```

The previous code is great but we don't have as much control as we'd like on where `Cache` is storing cached objects. To do that, we can explicitly provide a cache folder and add tags to the object so that we can find it more easily if we ever need to "clean it".
```{r LCC2005_4, eval = FALSE}
cPath <- file.path(tempdir(), "cache")
## run this twice
LCC2005 <- Cache(prepInputs,
                 url = url,
                 destinationPath = asPath(dPath),
                 useSAcrs = TRUE,  
                 overwrite = TRUE, 
                 studyArea = StudyArea,
                 cacheRepo = cPath,
                 userTags = "LCC2005")

showCache(x = cPath, userTags = "LCC2005")
reproducible::clearCache(x = cPath, userTags = "LCC2005")

## notice how Cache needs to re-do things
LCC2005 <- Cache(prepInputs,
                 url = url,
                 destinationPath = asPath(dPath),
                 useSAcrs = TRUE,  
                 overwrite = TRUE, 
                 studyArea = StudyArea,
                 cacheRepo = cPath,
                 userTags = "LCC2005")
```

We can also force `Cache` to redo operations and re-cache, or simply to ignore caching altogether. See more options for `Cache(useCache)` in `?Cache`
```{r LCC2005_5, eval = FALSE}
LCC2005 <- Cache(prepInputs,
                 url = url,
                 destinationPath = asPath(dPath),
                 useSAcrs = TRUE,
                 overwrite = TRUE,
                 studyArea = StudyArea,
                 cacheRepo = cPath,
                 userTags = "LCC2005",
                 useCache = "overwrite")

LCC2005 <- Cache(prepInputs,
                 url = url,
                 destinationPath = asPath(dPath),
                 useSAcrs = TRUE,
                 overwrite = TRUE,
                 studyArea = StudyArea,
                 cacheRepo = cPath,
                 userTags = "LCC2005",
                 useCache = FALSE)
```

Try to change the study area now. Hints:
* check out `?SpaDES.tools::randomPolygon`
  + try these coordinates: -114.5522 W 50.73892 N and 1 ha
* check out `?SpaDES.tools::randomStudyArea`
  + try with an area 1 ha
  + how would you guarantee you get *the same* area *everytime*?
* what happens when you run `Cache(prepInputs(...))` with the new study area(s)?
 
```{r LCC2005_6, eval = FALSE, include = FALSE}
library(SpaDES.tools)
StudyArea2 <- randomPolygon(x = matrix(c(-114.5522, 50.73892), nrow = 1, ncol = 2,
                                      byrow = TRUE), area = 100^2)
StudyArea3 <- randomStudyArea(seed = 1234, size = 100^2)

LCC2005 <- Cache(prepInputs,
                 url = url,
                 destinationPath = asPath(dPath),
                 useSAcrs = TRUE,
                 overwrite = TRUE,
                 studyArea = StudyArea2,
                 cacheRepo = cPath,
                 userTags = "LCC2005",
                 useCache = FALSE)
```

What if my study area is a raster?
Assuming you don't have raster at hand, try using a raster from the `SpaDESInAction` example ("inputs/LCC2005_V1_4a_BCR6_NWT.tif").

```{r LCC2005_7, eval = FALSE, include = FALSE}
templateRaster <- raster("docs/SpaDESInAction2/inputs/LCC2005_V1_4a_BCR6_NWT.tif")

LCC2005 <- Cache(prepInputs,
                 url = url,
                 destinationPath = asPath(dPath),
                 # useSAcrs = TRUE,
                 mask
                 overwrite = TRUE, 
                 rasterToMatch = templateRaster,
                 maskWithRTM = TRUE,
                 cacheRepo = cPath,
                 userTags = "LCC2005",
                 useCache = FALSE)
```


Learn more about [caching](http://spades.predictiveecology.org/articles/iii-cache.html)

# Some basic Debugging lessons

Now that you have a flavour of caching, we're going to explore debugging a bit and put our new "caching skills" in practice in a SpaDES modelling context.

We're going to run the `caribouRSF` module by itself.

```{r caribouRun, eval = FALSE}
library("reproducible")
library(SpaDES)
library(LandR)
library(raster)
library(data.table)

options(
  "spades.recoveryMode" = 2,
  "spades.lowMemory" = TRUE,
  "LandR.assertions" = FALSE,
  "LandR.verbose" = 1,
  "reproducible.useMemoise" = TRUE, # Brings cached stuff to memory during the second run
  "reproducible.useNewDigestAlgorithm" = TRUE,  # use the new less strict hashing algo
  "reproducible.useCache" = TRUE,
  "pemisc.useParallel" = FALSE
)

inputDirectory <- checkPath(file.path(getwd(), "inputs"), create = TRUE)
outputDirectory <- checkPath(file.path(getwd(), "outputs"), create = TRUE)
modulesDirectory <- checkPath(file.path(getwd(), "modules"), create = TRUE)
cacheDirectory <- checkPath(file.path(getwd(), "cache"), create = TRUE)

setPaths(cachePath = cacheDirectory,
         modulePath = c(modulesDirectory, 
                        file.path(modulesDirectory, "scfm/modules")),
         inputPath = inputDirectory,
         outputPath = outputDirectory)

times <- list(start = 0, end = 10)

successionTimestep <- 1L
parameters <- list(
  caribouRSF = list(
    "decidousSp" = c("Betu_Pap", "Popu_Tre", "Popu_Bal"),
    "predictionInterval" = 20,
  )
)
# load studyArea
studyArea <- readRDS(file.path(getPaths()$inputPath, "studyArea.rds"))

objects <- list(
  "studyArea" = studyArea
)

caribou <- simInitAndSpades(times = times,
                            objects = objects,
                            params = parameters,
                            modules = as.list("caribouRSF"),
                            paths = getPaths(),
                            debug = 1)
```

Oops, something doens't seem to be right! We start by looking carefully at the printed output, then we use `traceback` to help us locate the problem.
In this case, it seems to be a particular line of *caribouRSF.R*

```{r goingIn, eval=FALSE}
traceback()
# 11: stop("This module does not work without data. Please provide the necessary layers") at caribouRSF.R#154
# 10: get(moduleCall, envir = fnEnv)(sim, cur[["eventTime"]], cur[["eventType"]])
# 9: eval(fnCallAsExpr)
# 8: eval(fnCallAsExpr)
# (...)

file.edit("modules/caribouRSF/caribouRSF.R")  ## got to caribouRSF.R#154
```

1. Insert a `browser()` before the line with the `stop()`. Save and re-rerun.
+ Check `?browser`, while you're at it ;)
2. What is `P(sim)$.useDummyData`? Where does its value come from?
3. Which data objects are missing? Why?

```{r goingIn_2, eval = FALSE}
# Browse[1]> P(sim)$.useDummyData
# [1] TRUE

# Browse[1]> mod$pixelGroupMap
# NULL

# Browse[1]> mod$cohortData
# NULL
```

We are going to supply these objects - note that the dynamic part will not be simulated.

4. Check the `.inputObjects` function and the metadata for inputs.
+ Can you see a pattern in how `prepInputs` gets data from online sources? Try to do the same for `pixelGroupMap`and `cohortData`
+ How are sources for objects given? Try adding sources the following sources:
- for `pixelGroupMap`: "https://drive.google.com/open?id=1IUEuH55su8X7JCWt8LXy_hTAQz0cfCmU"
- for `cohortData`: [https://drive.google.com/open?id=1R_wGGvzUI0gGZ5NOs2KmT2KrmXaTm4NS]

```{r goingIn_3, include=FALSE, eval = FALSE}
## add sourceURL to pixelGroupMap and cohortData
expectsInput(objectName = "pixelGroupMap", objectClass = "RasterLayer",
             desc = paste0("Map of groups of pixels that share the same info from cohortData (sp, age, biomass, etc).",
                           "Here is mainly used to determine old and recent burns based on tree age,",
                           " and if deciduous by species"),
             sourceURL = "https://drive.google.com/open?id=1IUEuH55su8X7JCWt8LXy_hTAQz0cfCmU")
expectsInput(objectName = "cohortData", objectClass = "data.table",
             desc = paste0("data.table with information by pixel group of sp, age, biomass, etc"),
             sourceURL = "https://drive.google.com/open?id=1R_wGGvzUI0gGZ5NOs2KmT2KrmXaTm4NS")

## add defaults for these objects in .inputObjects, so that the module can get them if they are not supplied
if (!suppliedElsewhere("pixelGroupMap", sim = sim, where = "sim")) {
  sim$pixelGroupMap <- Cache(prepInputs, targetFile = "pixelGroupMapCaribouEg.rds",
                             fun = "readRDS",
                             url = extractURL("pixelGroupMap"), studyArea = sim$studyArea,
                             destinationPath = dataPath(sim), filename2 = NULL,
                             rasterToMatch = sim$rasterToMatch)
}

if (!suppliedElsewhere("cohortData", sim = sim, where = "sim")) {
  sim$cohortData <- Cache(prepInputs, targetFile = "cohortDataCaribouEg.rds",
                          fun = "readRDS",
                          url = extractURL("cohortData"),
                          destinationPath = dataPath(sim))
}
```

Learn about [debugging in SpaDES](https://github.com/PredictiveEcology/SpaDES/wiki/Debugging) and [with RStudio](https://support.rstudio.com/hc/en-us/articles/205612627-Debugging-with-RStudio)
