---
title       : "Caching and debugging"
author      : "Ceres Barros"
date        : "January 2020"
output: 
  html_document: 
    number_sections: yes
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, eval = FALSE)
```

# Some basic Caching lessons

Caching is the ability to save some sort of output from an operation, and then retrieve these outputs when the operation is repeated in the *same way* - meaning the inputs of this operation and the actual tasks it performs are unchanged.

Caching becomes fundamental when we can expect to re-run operations several times, particularly if they they a while to compute each time. Some examples of these operations are:
- downloading data
- (spatial) data processing/munging
- fitting statistical models to large datasets, or that are complex in nature
- running simulations with no stochasticity

`SpaDES` (via the `reproducible` package) offers a number of functions that make caching these operations a lot easier for non-programmers. Two fundamental ones are `Cache` and `prepInputs`.

```{r landCover_1}
library(reproducible)
library(sp)
library(raster)

## from ?prepInputs
dPath <- file.path("modules/Biomass_borealDataPrep/data")
url <- file.path("ftp://ftp.ccrs.nrcan.gc.ca/ad/NLCCLandCover",
                 "LandcoverCanada2005_250m/LandCoverOfCanada2005_V1_4.zip")
landCover <- prepInputs(url = url,
                        destinationPath = asPath(dPath))
```

Now do it again. Notice any difference?
```{r landCover_2}
landCover <- prepInputs(url = url,
                        destinationPath = asPath(dPath))
```

Now try wrapping the previous operation in `Cache` call, and run it twice. Notice differences in speed.
```{r landCover_3}
landCover <- Cache(prepInputs,
                   url = url,
                   destinationPath = asPath(dPath))


landCover <- Cache(prepInputs,
                   url = url,
                   destinationPath = asPath(dPath))
```

The previous code is great but we don't have as much control as we'd like on where `Cache` is storing cached objects. To do that, we can explicitly provide a cache folder and add tags to the object so that we can find it more easily if we ever need to "clean it".
```{r landCover_4}
cPath <- file.path(tempdir(), "cache")
## run this twice
landCover <- Cache(prepInputs,
                   url = url,
                   destinationPath = asPath(dPath),
                   cacheRepo = cPath,
                   userTags = "landCover")

showCache(x = cPath, userTags = "landCover")
reproducible::clearCache(x = cPath, userTags = "landCover")

## notice how Cache needs to re-do things
landCover <- Cache(prepInputs,
                   url = url,
                   destinationPath = asPath(dPath),
                   userTags = "landCover")
```

We can also force `Cache` to redo operations and re-cache, or simply to ignore caching altogether. See more options for `Cache(useCache)` in `?Cache`
```{r landCover_5}
landCover <- Cache(prepInputs,
                   url = url,
                   destinationPath = asPath(dPath),
                   cacheRepo = cPath,
                   userTags = "landCover",
                   useCache = "overwrite")
options("reproducible.useCache" = FALSE)
landCover <- Cache(prepInputs,
                   url = url,
                   destinationPath = asPath(dPath),
                   cacheRepo = cPath,
                   userTags = "landCover",
                   useCache = TRUE)
```

Try to provide a study area now. Hints:

+ check out `?reproducible::prepInputs` and `?reproducible::postProcess`
+ check out `?SpaDES.tools::randomStudyArea`
- try with an area 1 ha
+ what happens when you run `Cache(prepInputs(...))` with the new study area(s)?

```{r landCover_6}
library(SpaDES.tools)
StudyArea <- randomStudyArea(size = 100000^2)

## cheating to visualise beforehand
if (identical(crs(StudyArea), crs(landCover)))
  StudyArea <- spTransform(StudyArea, crs = crs(landCover))
plot(landCover)
plot(StudyArea, add = TRUE, col = "red")

options(reproducible.useCache = TRUE)
landCover <- Cache(prepInputs,
                   url = url,
                   destinationPath = asPath(dPath),
                   useSAcrs = TRUE,
                   overwrite = TRUE,
                   studyArea = StudyArea,
                   cacheRepo = cPath,
                   userTags = "landCover")
plot(landCover)
```

What if my study area is a raster?
Assuming you don't have raster at hand, try using a raster from the `SpaDESInAction` example ("inputs/rasterToMatch.rds").

```{r landCover_7}
## assuming you're in the SpaDESInAction2 project folder
templateRaster <- readRDS(file.path(getwd(), "inputs/rasterToMatch.rds"))

landCover <- Cache(prepInputs,
                   url = url,
                   destinationPath = asPath(dPath),
                   # useSAcrs = TRUE,  ## we don't need this anymore
                   rasterToMatch = templateRaster, ## a reproducible::postProcess argument
                   maskWithRTM = TRUE,  ## a reproducible::maskInputs argument 
                   overwrite = TRUE, 
                   cacheRepo = cPath,
                   userTags = "landCover",
                   useCache = TRUE)
plot(landCover)
```

What if I have both?
In some cases your study area may be defined by a polygon, but you may have a raster that will dictate the e.g. projection and resolution of the output (remember, polygons have no resolution).
```{r landCover_8}
landCover <- Cache(prepInputs,
                   url = url,
                   destinationPath = asPath(dPath),
                   studyArea = StudyArea,
                   useSAcrs = FALSE,  ## use the template raster projection
                   rasterToMatch = templateRaster,
                   maskWithRTM = FALSE,  ## mask using the study area
                   overwrite = TRUE, 
                   cacheRepo = cPath,
                   userTags = "landCover",
                   useCache = TRUE)
```

Now imagine someone told you there is a more up to date Land Cover map for Canada. And they told you where to look to get to the `.zip` file - [the Canadian Gov. open data portal](https://open.canada.ca/data/en/dataset/c688b87f-e85f-4842-b0e1-a8f79ebf1133)

+ Try right-click on "Access" for the TIF file and replace the "old" URL
+ Had an error, the messages are helpful!

```{r landCover_9}
url <- "http://ftp.maps.canada.ca/pub/nrcan_rncan/Land-cover_Couverture-du-sol/canada-landcover_canada-couverture-du-sol/CanadaLandcover2010.zip"
landCover <- Cache(prepInputs,
                   targetFile = "CAN_LC_2010_CAL.tif",
                   url = url,
                   destinationPath = asPath(dPath),
                   studyArea = StudyArea,
                   overwrite = TRUE, 
                   cacheRepo = cPath,
                   userTags = "landCover",
                   useCache = TRUE)
plot(landCover)
```

Learn more about [caching](http://spades.predictiveecology.org/articles/iii-cache.html)

# Some basic Debugging lessons

Now that you have a flavour of caching, we're going to explore debugging a bit and put our new "caching skills" in practice in a SpaDES modelling context.

We're going to run the `caribouRSF` module by itself.

```{r caribouRun, eval = FALSE}
## Restart your R session if not running from a "clean" environment
library("reproducible")
library(SpaDES)
library(LandR)
library(raster)
library(data.table)

options(
  "spades.recoveryMode" = 2,
  "spades.lowMemory" = TRUE,
  "LandR.assertions" = FALSE,
  "LandR.verbose" = 1,
  "reproducible.useMemoise" = TRUE, # Brings cached stuff to memory during the second run
  "reproducible.useNewDigestAlgorithm" = TRUE,  # use the new less strict hashing algo
  "reproducible.useCache" = TRUE,
  "pemisc.useParallel" = FALSE
)

## assuming you're in the SpaDESInAction2 project folder
inputDirectory <- checkPath(file.path(getwd(), "inputs"), create = TRUE)
outputDirectory <- checkPath(file.path(getwd(), "outputs"), create = TRUE)
modulesDirectory <- checkPath(file.path(getwd(), "modules"), create = TRUE)
cacheDirectory <- checkPath(file.path(getwd(), "cache"), create = TRUE)

setPaths(cachePath = cacheDirectory,
         modulePath = c(modulesDirectory, 
                        file.path(modulesDirectory, "scfm/modules")),
         inputPath = inputDirectory,
         outputPath = outputDirectory)

times <- list(start = 0, end = 10)

successionTimestep <- 1L
parameters <- list(
  caribouRSF = list(
    "decidousSp" = c("Betu_Pap", "Popu_Tre", "Popu_Bal"),
    "predictionInterval" = 20
  )
)
# load studyArea
studyArea <- readRDS(file.path(getPaths()$inputPath, "studyArea.rds"))

objects <- list(
  "studyArea" = studyArea
)

caribou <- simInitAndSpades(times = times,
                            objects = objects,
                            params = parameters,
                            modules = as.list("caribouRSF"),
                            paths = getPaths(),
                            debug = 1)
```

Oops, something doesn't seem to be right! We start by looking carefully at the printed output, then we use `traceback` to help us locate the problem.
In this case, it seems to be a particular line of *caribouRSF.R*

```{r goingIn, eval=FALSE}
traceback()
# 11: stop("This module does not work without data. Please provide the necessary layers") at caribouRSF.R#154
# 10: get(moduleCall, envir = fnEnv)(sim, cur[["eventTime"]], cur[["eventType"]])
# 9: eval(fnCallAsExpr)
# 8: eval(fnCallAsExpr)
# (...)

file.edit("modules/caribouRSF/caribouRSF.R")  ## got to caribouRSF.R#154
```

1. Insert a `browser()` before the line with the `stop()`. Save and re-rerun.
OR
2. Use the `debug` option in `simInitAndSpaDES` or `spades` to go in into "browser mode"

+ Check `?browser`, while you're at it ;)

2. What is `P(sim)$.useDummyData`? Where does its value come from?
3. Which data objects are missing? Why?

```{r goingIn_2, eval = FALSE}
# Browse[1]> P(sim)$.useDummyData
# [1] TRUE

# Browse[1]> mod$pixelGroupMap
# NULL

# Browse[1]> mod$cohortData
# NULL
```

You can also try to enter the "browser mode" in specific events or all events of a module via the `simInitAndSpades` or `spades` functions.

```{r goingIn_3, eval = FALSE}
## browse at the init event(s) - if you had more than one module it would stop at each
caribou <- simInitAndSpades(times = times,
                            objects = objects,
                            params = parameters,
                            modules = as.list("caribouRSF"),
                            paths = getPaths(),
                            debug = "init")

## browse at the event that triggered the error
caribou <- simInitAndSpades(times = times,
                            objects = objects,
                            params = parameters,
                            modules = as.list("caribouRSF"),
                            paths = getPaths(),
                            debug = "lookingForCaribou")

## "browse" at each event of the caribouRSF module
caribou <- simInitAndSpades(times = times,
                            objects = objects,
                            params = parameters,
                            modules = as.list("caribouRSF"),
                            paths = getPaths(),
                            debug = "caribouRSF")
```


We are going to supply these objects - note that the dynamic part will not be simulated.

4. Check the `.inputObjects` function and the metadata for inputs.

+ Can you see a pattern in how `prepInputs` gets data from online sources? Try to do the same for `pixelGroupMap`and `cohortData`
+ How are sources for objects given? Try adding sources the following sources:

- for `pixelGroupMap`: "https://drive.google.com/open?id=1IUEuH55su8X7JCWt8LXy_hTAQz0cfCmU"
- for `cohortData`: "https://drive.google.com/open?id=1R_wGGvzUI0gGZ5NOs2KmT2KrmXaTm4NS"

```{r goingIn_4, eval = FALSE}
## add sourceURL to pixelGroupMap and cohortData
expectsInput(objectName = "pixelGroupMap", objectClass = "RasterLayer",
             desc = paste0("Map of groups of pixels that share the same info from cohortData (sp, age, biomass, etc).",
                           "Here is mainly used to determine old and recent burns based on tree age,",
                           " and if deciduous by species"),
             sourceURL = "https://drive.google.com/open?id=1IUEuH55su8X7JCWt8LXy_hTAQz0cfCmU")
expectsInput(objectName = "cohortData", objectClass = "data.table",
             desc = paste0("data.table with information by pixel group of sp, age, biomass, etc"),
             sourceURL = "https://drive.google.com/open?id=1R_wGGvzUI0gGZ5NOs2KmT2KrmXaTm4NS")

## add defaults for these objects in .inputObjects, so that the module can get them if they are not supplied
if (!suppliedElsewhere("pixelGroupMap", sim = sim, where = "sim")) {
  sim$pixelGroupMap <- Cache(prepInputs, targetFile = "pixelGroupMapCaribouEg.rds",
                             fun = "readRDS",
                             url = extractURL("pixelGroupMap"), studyArea = sim$studyArea,
                             destinationPath = dataPath(sim), filename2 = NULL,
                             rasterToMatch = sim$rasterToMatch)
}

if (!suppliedElsewhere("cohortData", sim = sim, where = "sim")) {
  sim$cohortData <- Cache(prepInputs, targetFile = "cohortDataCaribouEg.rds",
                          fun = "readRDS",
                          url = extractURL("cohortData"),
                          destinationPath = dataPath(sim))
}
```

You can now use `restartSpades` or simply re-run `simInitAndSpades`
```{r restartSpades, eval = FALSE}
caribou <- restartSpaDES()

caribou <- simInitAndSpades(times = times,
                            objects = objects,
                            params = parameters,
                            modules = as.list("caribouRSF"),
                            paths = getPaths(),
                            debug = 1)
```

There seem to be no more issues. Maybe we don't need to print all those debug messages anymore - less verbose - next time we run.

```{r debug1, eval = FALSE}
caribou <- simInitAndSpades(times = times,
                            objects = objects,
                            params = parameters,
                            modules = as.list("caribouRSF"),
                            paths = getPaths(),
                            debug = FALSE)
```

Or maybe, we don't want to see them printed, but want to keep a log of all messages to check later.

+ check the "debug" section of `?spades` 

```{r debug2, eval = FALSE}
caribou <- simInitAndSpades(times = times,
                            objects = objects,
                            params = parameters,
                            modules = as.list("caribouRSF"),
                            paths = getPaths(),
                            debug = list(console = list(level = 40), 
                                         file = list(append = FALSE,
                                                     file = "logCaribou.txt",
                                                     level = 0), 
                                         debug = TRUE)
)

## notivce how different the file is from what was printed on the console.
file.edit("logCaribou.txt")
```

What about all that purple text?
That's the module code checking. It's helpful, but not always accurate - meant to be informative rather than enforced.
```{r debug3, eval = FALSE}
options("spades.moduleCodeChecks" = FALSE)
caribou <- simInitAndSpades(times = times,
                            objects = objects,
                            params = parameters,
                            modules = as.list("caribouRSF"),
                            paths = getPaths(),
                            debug = list(console = list(level = 40), 
                                         file = list(append = FALSE,
                                                     file = "logCaribou.txt",
                                                     level = 0), 
                                         debug = TRUE)
)

## notice how different the file is from what was printed on the console.
file.edit("logCaribou.txt")
```

Learn about [debugging in SpaDES](https://github.com/PredictiveEcology/SpaDES/wiki/Debugging) and [with RStudio](https://support.rstudio.com/hc/en-us/articles/205612627-Debugging-with-RStudio)
