---
title       : "The 'data-to-decisions' workflow"
author      : "Alex M Chubaty & Eliot McIntire"
date        : "December 9, 2016"
output:
  ioslides_presentation:
    logo: images/predictive_ecology_logo.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, eval = FALSE)
```

## Reproducible pipeline

<div class="centered" style="color:#0000FF">
*…This places an obligation on all creators of software to program in such a way that the computations can be understood and trusted. This obligation I label the Prime Directive.*
</div>

- John Chambers (Software for Data Analysis: Programming with R)

## Adaptive management

- Ideally, we can rerun all decision support tools over and over again
- Must be MUCH easier the second, third, etc. time
- If there is an incremental change in input data, then we shouldn't need to redo the same effort as the first time around.

## Why a Discrete Event Simulator is necessary?

- There is much fanfare about using Rmarkdown as a tool for reproducibility (*i.e*, Rmarkdown = "code, output, narrative")
- But this allows *exact* reproducibility

    - **Everything will be run exactly as before**
    
- What happens if the goal is to run it again, but with slight changes to the input data?
- The Rmarkdown approach can deal with this to a certain extent, but it will be clunky

## Adding novelty

- What happens if you want to add something new, but keep it mostly the same
- You have a good vegetation model and a good caribou model, but you want to add a wolf model
- With Rmarkdown, you would have to edit the original file
- With `SpaDES`, you can use the other modules by calling them... no need to know how they work
- You add your own wolf model, looking at outputs of the vegetation model and caribou model
- Using the model metadata from those other modules

## Deploying for policy and management

Need the reusable work flow, and reproducible science ... but with a GUI

- `shine` function is a start
- we created a demo version of a generic web interface for models ([link](http://predictiveecology.org/2016/11/17/Putting_science_in_the_hands_of_policy_makers.html))
- A non-scientist can ask questions of a system of modules without knowing which modules are necessary

## Caching

- For the above to work for, say, a national question, it can't run the model on the fly
- Could have massive data requirements, long simulation times, even on supercomputers
- Caching allows for near instant results from very complex systems

```{r caching for data2decisions}
?Cache
```

[wolf model shows a caching use case](http://htmlpreview.github.io/?https://github.com/PredictiveEcology/SpaDES-modules/blob/master/modules/wolfAlps/wolfAlps.html)

## Data-to-decisions

- This is the current vision for `SpaDES`
- We are looking for people who are interested in being part of this adventure

<div class="centered" style="color:#0000FF">
*Bringing the best science, data, models into the hands of policy makers in real time, on their phones*
</div>
