---
title       : "R -- developing powerful skills"
author      : "Alex M Chubaty & Eliot McIntire"
date        : "December 9, 2016"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=FALSE, echo=TRUE, eval=FALSE)
```

## Need for speed

R is widely seen as being 'slow' (see [julia](http://julialang.org/#high-performance-jit-compiler) web page) 

**But**, if you use a few specific tools, then this becomes irrelevant because of the powerful tools in various packages in R
 
## An aside
 
Pure R, when the most efficient vectorized code is used, appears to be 1/2x the speed of the most efficient C++.

See Hadley Wickham's [page on Rcpp](http://adv-r.had.co.nz/Rcpp.html), scroll down to "Vector input, vector output"... ), noting that if it took 10 minutes to write the C++ code, it would have to be 150,000 times faster to make it worth it.
 
## Need for speed

Spatial simulation means doing the same thing over and over and over ... so we need speed

We will show how to profile your code at the end of this section.

## Predictive Ecology blog posts about R speed

- [Is R Fast Enough, part 1, The Mean](http://predictiveecology.org/2015/04/23/Is-R-fast-enough-01.html)
- [Is R Fast Enough, part 2, Sorting](http://predictiveecology.org/2015/04/28/Is-R-fast-enough-02.html)
- [Is R Fast Enough, part 3, Fibonacci](http://predictiveecology.org/2015/05/06/Is-R-fast-enough-03.html)
- [Is R Fast Enough, part 4, Loops](http://predictiveecology.org/2015/09/04/Is-R-fast-enough-04.html)

## "Vectorization"

- This is at the core of making R fast. If you don't do this, then it is probably not useful to use R as a simulation engine.

```{r vectorization}
# Instead of 
a <- vector()
for (i in 1:1000) {
  a[i] <- rnorm(1)
}

# use vectorized version, which is built into the functions
a <- rnorm(1000)
```

## Vectors and Matrices

- These are as fast as you can get in R
- Fast numerical operations
- Faster than `data.frame`
- Anything that is in pure vectors or matrices is 'fast enough'
- It is always a challenge to keep all code in vectors and matrices
- Thus the following packages...

## Spatial simulation

- To work with spatial simulation (*e.g.*, time and space), it requires more than *just* spatial data manipulation
- Sometimes it is just base R stuff
- Need to learn how to make functions (allows reusability)
- Need to learn a few key packages that are critical for speed

## Key packages for spatial simulation

- `base` package -- everything matrix or vector is 'fast'
- `raster` - for spatially referenced matrices

    - not always fast enough, sometimes we copy the data into a matrix, then manipulate, then return the data to the raster object
    
- `sp` - equivalent of vector shapefiles in a GIS

    - Polygons, Points, Lines
    - Not always fast, but essential to have

- see also `sf`

## Key packages for spatial simulation
    
- `data.table`

    - For `data.frame` type data (i.e., columns of data)
    - Very fast when object gets large, but is actually slower if the `data.frame` is small (<100,000 rows)
    
- `SpaDES` -- many functions; will be moved into a separate package soon

- `Rcpp`

    - R interface to C++ . When you need something fast, and you can't get it fast enough with existing tools/packages, you can create your own (we will not go further into this here)

## What we will do here

- We will go through `SpaDES` functions quickly, because there are fewer tutorials online for these
- We will show links to various tutorials for `raster`, `sp`, `data.table`, `Rcpp`
- Each person should decide which tool is the most useful to them
- Put something into practice

## `SpaDES` functions

- These are all potentially useful for building spatio-temporal models
```{r SpaDES functions}

?`spades-package` # section 2 shows many functions

# e.g.,
?spread
?move
?cir
?adj
?distanceFromEachPoint
```

## Working with spatial data

- [Excellent Cheatsheet](http://www.maths.lancs.ac.uk/~rowlings/Teaching/UseR2012/cheatsheet.html)

## `raster`

- tutorials 

    - [NEON](http://neondataskills.org/R/Raster-Data-In-R/)
    - [geoscripting-wur](http://geoscripting-wur.github.io/IntroToRaster/)
    - [Official vignette](https://cran.r-project.org/web/packages/raster/vignettes/Raster.pdf)
    - [Using R as a GIS](https://pakillo.github.io/R-GIS-tutorial/)
    - many more available

## `sp`

- Quite an old and mature package
- Tutorials

    - [stack exchange](http://gis.stackexchange.com/questions/45327/tutorials-to-handle-spatial-data-in-r)
    - [NEON](http://neondataskills.org/R-package/sp/)
    

## `sf`

- Relatively new
- Implements latest GIS data standards
- Very fast, especially reading/writing large data
- [CRAN](https://cran.r-project.org/web/packages/sf/)
- [GitHub](https://github.com/edzer/sfr)

## The `data.table` package

From every `data.table` user ever:

<div class="centered" style="color:#0000FF">*WOW that's fast!*</div>

<br>

```{r install-data.table, echo=TRUE, eval=FALSE}
install.packages('data.table')
```

(at least for large tables!)

## `data.table` tutorials

- [Data Camp](https://www.datacamp.com/community/tutorials/data-table-r-tutorial#gs.s=AE7o0)
- [Cheat sheet](https://www.datacamp.com/community/tutorials/data-table-cheat-sheet#gs.rFn247Q)

## `raster` and `data.table` together

- The current implementation of `LANDIS-SpaDES` uses a "reduced" data structure throughout

- Instead of keeping rasters of everything (one can imagine that there is redundancy, i.e., 2 pixels next to each other may be identical)

- We make one raster of "id" and one data.table with a column called "id"

- Then we can have as many columns as we want of information about each of these places

- Like "polygons", but for rasters, and dynamic... can change over time

- This may be useful for your own module

## `raster` and `data.table` together

- There is a key helper function:

```{r datatable and raster}
?rasterizeReduced
```

What does this do?

## The `Rcpp` package

From every `Rcpp` user ever:

<div class="centered" style="color:#0000FF">*WOW! Just wow.*</div>

<br>

```{r install-rcpp, echo=TRUE, eval=FALSE}
install.packages('Rcpp')
```

- Essentially see [Dirk Eddelbuettel's book, Seamless R and C++](http://dirk.eddelbuettel.com/code/rcpp.html)
- [Hadley's section in Advanced R](http://adv-r.had.co.nz/Rcpp.html)

