---
title       : "R -- Profiling code"
author      : "Alex M Chubaty & Eliot McIntire"
date        : "December 9, 2016"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=FALSE, echo=TRUE, eval=FALSE)
```

## Profiling and Benchmarking (1)

- In general, the usual claim is to worry about 'execution speed later'
- This is not 100% true with R
- If you use vectorization (no or few loops), and these packages listed here, then you will have a good start
- AFTER that, then you can use 2 great tools:

    - `profvis` package (built into the latest Rstudio previews, but not the official release version)
    - `microbenchmark` package

## Profiling and Benchmarking (2)

```{r microbenchmark, eval=TRUE, echo=TRUE}
microbenchmark::microbenchmark(
  loop = {
    a <- vector()
    for (i in 1:1000) a[i] <- runif(1)
  },
  vectorized = { a <- runif(1000) }
)
```

## Profiling and Benchmarking (3)

If you have Rstudio version `>=0.99.1208`, then it has profiling as a menu item.

- alternatively, we wrap any block of code with `profvis`

- This can be a `spades()` call, so it will show you the entire model:

```{r profiling, eval=FALSE, echo=TRUE}
profvis::profvis({a <- rnorm(10000000)})
```

## Profiling the `spades` call

Try it:

```{r profiling spades, eval=FALSE, echo=TRUE}
mySim <- simInit(
   times = list(start = 0.0, end = 2.0, timeunit = "year"),
   params = list(
     .globals = list(stackName = "landscape", burnStats = "nPixelsBurned")
   ),
   modules = list("randomLandscapes", "fireSpread", "caribouMovement"),
   paths = list(modulePath = system.file("sampleModules", package = "SpaDES"))
)
profvis::profvis({spades(mySim)})
```

## When to profile

- First, you should have started building your code with the packages we have discussed
- It will be too late if you have loops in your code, and you are ready to profile to improve it

If you have used these tools, then:

- When you have mostly finished whatever we are coding
- Don't ever start making code more efficient until you have profiled
- It is almost impossible to tell which bits are the slow parts, without profiling or benchmarking

## Strategies for profiling

- Can do an entire `SpaDES` model call
- Can pinpoint specific functions
- Can test alternative ways of implementing the same thing
