---
title       : "Sharing modules and data"
author      : "Alex M Chubaty & Eliot McIntire"
date        : "December 9, 2016"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, eval = FALSE)
library(SpaDES)
```

## `SpaDES` module repositories

A 'module repository' is simply a web site that serves files using a particular directory structure (the same directory structure used in a local `SpaDES` module repo).

Our module repo is at https://github.com/PredictiveEcology/SpaDES-modules

This is the default repo in `SpaDES`.

    - it's public
    - it's easily accessible using `downloadModule()`

```{r module-repo-option, echo=TRUE, eval=TRUE}
getOption('spades.modulesRepo') ## default url prefix is GitHub
```

## Other `SpaDES` module repositories

Anybody can create their own repository as well (R has CRAN, Rforge, BioConductor and many lesser used ones)

`downloadModule()` has a `repo` argument.

## Our `SpaDES` module repository

- Initially, most `SpaDES` modules will not be there
- *e.g.*, SpaDES-LBM (SpaDES -- Landis Biomass Succession) is not there yet (we would like to publish it first)

## What is GitHub?

[Github.com](https://github.com/) is a free\* code archive and hosting service.

Allows hosting of public and private code repositories, build around `git` version control software.

GitHub provides:

- code archiving/distribution
- version control
- publication (code, html pages, and others)
- collaboration
- bug tracker
- **user friendly web interface and desktop client**

## Using GitHub for collaboration

We, and many others, use GitHub extensively for all our group's work

- Like Dropbox, in the sense that there is a cloud version, and potentially many local copies all over the place
- Unlike Dropbox, the copies on people's computers are not automatically there
- Requires manual intervention

## GitHub Client

- For your local copy of the files, you need an extra piece of software on your computer (like the "Dropbox client")

- `git` and GitHub both require a bit of learning:

    - [tutorial](https://guides.github.com/activities/hello-world/)

- There are many other great `git` clients:

    - **[Gitkraken](http://gitkraken.com/)** is amazing and our group's prefered client (cross-platform)
    - [SourceTree](https://www.sourcetreeapp.com/) (Windows & Mac only)
    - See https://git-scm.com/downloads/guis for more

## `shiny` apps

- Massively powerful
- Allows the process of data wrangling, data visualization, data analysis, etc. to be made into interactive web (and mobile) apps
- By moving the web app development into R (instead of taking R outputs and putting them into a web development kit), means that the analysis power can be at the interactive stage
- Also, data analysts can be making web pages

## `shine` function

`SpaDES` has a simple function that takes any simList and makes a web app from it:

```{r shine}
?shine
```

<div class="centered">Try it!</div>

## Simple shiny apps

- hosted apps at [shinyapps.io](http://www.shinyapps.io/) 

    - *e.g.*, [Proof of concept](https://spades.shinyapps.io/ForestChange_ProofOfConcept/)

## Themes

- Lots of developers making shiny themes
- Often they are wrappers on javascript code
- So your look can be easily modified without many web development skills

## Shiny maps with leaflet

- [leaflet package](https://rstudio.github.io/leaflet/)

# Data sources

## Data sharing

- We have shown a way to specify default data sources

*Where was that?*

- We have shown a way to specify default data manipulations and data preparations of those default data

*Where was that?*

## If there are complex data sets?

- And from many sources?
- Can still use `sourceURL` in `defineModule` and create an `.inputObjects` function which will do all the manipulations

## How to adequately describe data?

- It is clear that there is not enough information contained in the metadata to fully describe everything about the required dataset(s) type(s)
- **Use the .Rmd** file to describe more for human eyes
- It is very difficult to algorithmically define data in a generic way
- There are standards, but `SpaDES` doesn't tap into those (yet!)

## Overriding default datasets

- If a *user* passes an object that a module requires (in 'defineModule' metadata)
- The default data will not be used

- checksums
- externally hosted data
- include data with your module code
