---
title       : "Simulation experiments and replication"
author      : "Alex M Chubaty & Eliot McIntire"
date        : "December 8, 2016"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, eval = FALSE)
```

## Beyond `spades`

So far, we have run models with the `spades` function.  

`spades()` only has a few arguments:

    - debug
    - .plotInitialTime
    - .saveInitialTime
    - progress
    - cache

What if we want to do more stuff?

## Replication

What if we have a stochastic model and need many independent (Monte Carlo) runs?

The `experiment` function does this.

```{r}
?experiment # for details and examples
```

Generally, however, this creates problems because of the computational time required ...

## Replication

The `replicates` argument

It is as easy as `replicates = 2` to do two replicates:

```{r experiment2}
?experiment # for details

# See Example 5
#  ... run mySim from top of help file
sims <- experiment(mySim, replicates = 2)
attr(sims, "experiment")$expDesign # shows 2 reps of same expt
```

What does it return?

What is the `attr`?

## The curse of replication

Replication is necessary but it multiplies the time it takes to run any model.

Fortunately, replication can use parallel processing:

- This is the simplest, yet most effective, parallel processing;
- No communication between individual simulations, until the end;
- So they can be spread across threads on a single computer or among computers;
- Requires a cluser (`cl`) argument.

## Parallel

- `SpaDES` functions that are parallel aware:

    - `experiment`, `POM`, `splitRaster`
    
- In each case, there are two ways:

```{r parallel}
## option 1:
raster::beginCluster() # make a cluster silently in background

## option 2:
cl <- parallel::makeCluster() # need to explicitly pass the
                              # cluster object into functions
                              # using cl argument
```

Try it!

## `experiment`: Varying parameter values

```{r experiment-params}
?experiment
## see example 1
experimentParams <- list(fireSpread = 
                           list(spreadprob = list(0.2, 0.23),
                                nFires = list(20, 10)),
                         caribouMovement = 
                           list(N = list(100, 1000)))
sims <- experiment(mySim, params = experimentParams)
attr(sims, "experiment")
```

## `experiment`: Saving files

Often, a `spades` call will generate output files.
Where do they go when using `experiment`?

```{r experiment-saving}
?experiment
## see example 4
sims <- experiment(mySim, 
                   params = experimentParams, 
                   dirPrefix = c("expt", "simNum"))
attr(sims, "experiment")$expVals # shows 8 alternative
               #experiment levels, 24 unique parameter values

dir(outputPath(mySim))
```

## Working through other examples of `experiment`

Exercise:

Pick a few examples and try to understand them.
(**NOTE:** they get more complicated as you get towards the bottom.)

```{r experiment-examples }
?experiment
```

## Running parallel simulations

Clearly, using `experiment` can take a lot of resources, but there are a few options:

- Use your own computer, a local cluster, a server, cloud and more;

- Access Compute Canada for free (though sometimes the queues are long):
    
    - [Blog post on using Compute Canada](http://predictiveecology.org/2016/07/04/How-to-start-using-a-HPC-cluster.html)
    
    - There is a different example here: [ssh among Linux machines](http://htmlpreview.github.io/?https://github.com/PredictiveEcology/SpaDES-modules/blob/master/modules/wolfAlps/wolfAlps.html#below-here-is-for-advanced-users)  
    
        - Shows how to use a Linux cluster connected via SSH;
        - Can be very effective because is interactive (unlike normal compute clusters).

# Pattern Oriented Modeling

## Estimating unknown parameters

- Ecological models use equations and functions
- These have parameters
- Parameters can be estimated from:

    - directly from data (*e.g.*, regression, machine learning)
    - indirectly using (*e.g.*, pattern oriented modeling) (*e.g.*, Grimm *et al.* Science 2005, Marucco & McIntire J Appl Ecol 2010)  

## Directly from data

- (Just a brief discussion of this here)
- Remember, we are using R!
- So you can use any statistical method that R has
- Mixed models, non-linear models, machine learning, bayesian

```{r direct from data}
nTrees <- 50
diamCM <- rlnorm(nTrees, meanlog = 2, sdlog = 0.7)
height <- rnorm(nTrees, sqrt(diamCM)*2 + rnorm(nTrees))
glmObj <- glm(height ~ diamCM)
plot(diamCM, height)
summary(glmObj)
```

## Predict method

- These methods often have predict methods associated with them
- So, instead of using "fixed" parameters in simulation models
- We can use these methods which give more accurate prediction:

    - given uncertainty, 
    - random effects, 
    - covariance matrix

```{r direct from data 2}
# predict with new data
predVals <- predict(glmObj, newdata = data.frame(
  diamCM = rlnorm(nTrees, meanlog = 2, sdlog = 0.7)
))
```

## Using this in simulation models

- In R, everything is an object
- The `glmObj` and `predictVals` are objects
- These have many features that can be used in simulations

    - plot methods
    - standard errors, covariance matrices
    - random effects terms
    
- For the first 25 years of simulation modeling in ecology, parameters were copied into simulation models and hard coded
- This made it difficult to update the simulation model if there are new data and the regression needs to be re-fit
- Using `SpaDES` we can now fit the data and use the parameters all inside the simulation

## Indirectly from data

### Pattern Oriented Modeling (POM)

- What does this mean?

## Heuristic Optimization

- Optimization is the process of minimizing some objective function
- Heuristic optimization is when there is no derivative-based solution
- Requires simulation-based approaches
- In R, currently, `DEoptim` seems to be the best optimization package for heuristic optimization
- `POM` function in `SpaDES` uses this package

## `POM`

There are several examples in `?POM`, which we will go through:

```{r POM1}
## example 1
# What is example 1 doing?
# Run it
# What does the result mean?
```

## More complicated `POM` 

- What would be next?

## More complicated `POM` - example 2

- What would be next?

```{r POM2}
## example 2
# What is example 2 doing?
# Run it
# What does the result mean?
```

## More complicated `POM` - example 3

- What would be next?

```{r POM3}
## example 3 
# What is example 3 doing?
# Run it
# What does the result mean?
```

*****

Often we want to run a simulation with a variety of parameter values as inputs. This is sometimes called a simulation experiment. The `experiment` function gives us an easy interface to this.

We will use the `establishment` module that we created in exercise 4a.

# Download a module

```{r experiment}
library(SpaDES)
module.path <- file.path(dirname(tempdir()), "modules")
downloadModule("establishment", path = module.path)
setPaths(modulePath = module.path)
library(ggplot2)

modules <- list("establishment")

mySim <- simInit(modules = modules)

# make sure the plotting device is clear
clearPlot()
spades(mySim)
```

# Create an experiment

1. Find out which parameters are in this module. (Hint: where is the module? Can you open it easily from R?) 
2. Pick one of the parameters and create a range of values for it. 
3. Following the structure indicated in example 4 of `?experiment`, use `experiment` to run a `spades` call for each parameter set (note that `experiment` is just a wrapper around `spades`). 
4. Inspect the experiment structure using `attributes(mySim)`. 
5. Create a plot showing on the x axis your parameter values, and on the y-axis something like `sum(mySim$establish[])`, *i.e.*, the number of pixels that had an establishment event. 

# Advanced - Create a two-parameter experiment

1. Repeat as above, but vary two parameters.

[See a solution here](https://github.com/PredictiveEcology/SpaDES-modules/blob/master/modules/establishment/establishment.Rmd)
