---
title       : "Getting technical"
author      : "Alex M Chubaty & Eliot McIntire"
date        : "December 7, 2016"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, eval = FALSE)
library(magrittr)
library(SpaDES)
```

## The parts

1. Initializing a simulation using `simInit()` creates a `simList` object.

2. Running a simulation via a `spades()` call returns a modified `simList` object.

## How are simulations specified?

A `simList` object is simply a structured data type containing various elements of a simulation.

The main components of a `simList` are:

1. A list of modules used;
2. The event queue;
3. A description of the data (object) dependencies.

## How `SpaDES` works: simLists

We can examine the `simList` object structure in the usual R fashion by printing (`show`ing) it, or by using `str()`:

```{r simlist-structure, echo=TRUE, eval=FALSE}
emptySim <- simInit()
emptySim  # same as show(emptySim)
str(emptySim)
```

**NOTE:** `simList`s are S4 objects, so we can use `getSlots()` and `slotNames()` to examine the object.

See also `?'.simList-class'`

**Exercises 03, numbers 2 & 3.**

## Accessing the parts of a simLists

<div class="columns-2">
```{r simList-accessors, echo=FALSE, eval=TRUE}
emptySim <- simInit()
slots <- paste0(slotNames(emptySim))
accessors <- c('envir', 'modules', 'params', 'events', 'current', 'completed', 'depends', 'times', 'inputs', 'outputs', 'paths') %>% 
  paste0('()')
data.frame(slot = slots, accessor = accessors) %>%
  knitr::kable(format = 'html', padding = 0)
```

```{r simList-accessors2, echo=FALSE, eval=TRUE}
other_accessors <- c('packages', 'globals', 'start', 'end', 'timeunit', 'timeunits', 'objects', 'paths', 'cachePath', 'inputPath', 'outputPath', 'modulePath') %>%
  paste0('()')
data.frame(other_accessors = other_accessors) %>%
  knitr::kable(format = 'html', padding = 0)
```
</div>

## How `SpaDES` works: `spades()`

### Model specification

Simple examples (using demo modules) of `simInit()` and `spades()` calls.

**Exercises 03, number 4, 5, & 6.**

- examine simList structure before and after the `spades()` call

## Using pre-built modules

1. Browse for modules in our `SpaDES` module repository:

    [https://github.com/PredictiveEcology/SpaDES-modules](https://github.com/PredictiveEcology/SpaDES-modules)

2. Download a module, its data, and required packages:

    ```{r downloadModule, eval=FALSE}
    module.path <- file.path(dirname(tempdir()), "modules")
    downloadModule('wolfAlps', module.path, data = FALSE)
    downloadData('wolfAlps', module.path)
    f <- file.path(module.path, 'wolfAlps', 'wolfAlps.R')
    pkgs <- packages(filename = f)
    install.packages(pkgs[-3]) ## NetLogoR not yet on CRAN
    ```

## Customizing an existing module

**Important:** always make a copy of the module you wish to modify before editing!

Open the module's code file in your editor:

```{r openModules, eval=FALSE}
openModules('wolfAlps', module.path)  # opens only the named module
openModules(path = module.path)       # opens all modules in a directory
```

## Next steps

Building `SpaDES` modules from scratch ([slides](http://rpubs.com/PredictiveEcology/SpaDES-Intro-04-modules)).
