---
title       : "simList -- What is in it?"
author      : "Alex M Chubaty & Eliot McIntire"
date        : "December 7, 2016"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, eval = FALSE)
library(magrittr)
library(SpaDES)
```

## Where does `simList` come from?

1. Initializing a simulation using `simInit()` creates a `simList` object.

2. Running a simulation via a `spades()` call returns a modified `simList` object.

## How are simulations specified?

A `simList` object is simply a structured data type containing various elements of a simulation.

The main components of a `simList` are:

1. A list of modules used;
2. The event queue;
3. A description of the data (object) dependencies.

## How `SpaDES` works: simLists

We can examine the `simList` object structure in the usual R fashion by printing (`show`ing) it, or by using `str()`:

```{r simlist-structure, echo=TRUE, eval=FALSE}
emptySim <- simInit()
emptySim  # same as show(emptySim)
str(emptySim)
```

**NOTE:** `simList`s are S4 objects, so we can use `getSlots()` and `slotNames()` to examine the object.

See also `?'.simList-class'`

## Accessing the parts of a simLists

<div class="columns-2">
```{r simList-accessors, echo=FALSE, eval=TRUE}
emptySim <- simInit()
slots <- paste0(slotNames(emptySim))
accessors <- c('envir', 'modules', 'params', 'events', 'current', 'completed', 'depends', 'times', 'inputs', 'outputs', 'paths') %>% 
  paste0('()')
data.frame(slot = slots, accessor = accessors) %>%
  knitr::kable(format = 'html', padding = 0)
```

```{r simList-accessors2, echo=FALSE, eval=TRUE}
other_accessors <- c('packages', 'globals', 'start', 'end', 'timeunit', 'timeunits', 'objects', 'paths', 'cachePath', 'inputPath', 'outputPath', 'modulePath') %>%
  paste0('()')
data.frame(other_accessors = other_accessors) %>%
  knitr::kable(format = 'html', padding = 0)
```
</div>


## Model specification

Simple examples (using demo modules) of `simInit()` and `spades()` calls.

- examine `simList` structure before and after the `spades()` call


### Initiliaze an empty `simList` object

```{r empty-simList}
emptySim <- simInit()
```

### Inspect the empty `simList` object

1. What are the names of each of the slots in the simList?
2. What do each of these slots contain? Hint: see `?'.simList-class'`.

```{r inspect-emptySim}
str(emptySim)
slotNames(emptySim)
getSlots('simList')
?'.simList-class'
```

### Initialize a simulation using sample modules

This code is based on the `SpaDES` package [demo](https://github.com/PredictiveEcology/SpaDES/blob/master/demo/spades-simulation.R).

```{r demo}
set.seed(42)

filelist <- data.frame(
  files = dir(system.file("maps", package = "SpaDES"),
              full.names = TRUE, pattern = "tif"),
  functions = "rasterToMemory",
  packages = "SpaDES",
  stringsAsFactors = FALSE
)

# create the simList object 'mySim'
mySim <- simInit(
  times = list(start = 0.0, end = 100.00),
  params = list(
    .progress = list(type = NA, interval = NA),
    .globals = list(burnStats = "nPixelsBurned"),
    randomLandscapes = list(
      nx = 1e2, ny = 1e2, .saveObjects = "landscape",
      .plotInitialTime = NA, .plotInterval = NA, inRAM = TRUE
    ),
    caribouMovement = list(
      N = 1e2, .saveObjects = c("caribou"),
      .plotInitialTime = 1, .plotInterval = 1, moveInterval = 1
    ),
    fireSpread = list(
      nFires = 1e1, spreadprob = 0.235, persistprob = 0, its = 1e6,
      returnInterval = 10, startTime = 0,
      .plotInitialTime = 0, .plotInterval = 10
    )
  ),
  modules = list("randomLandscapes", "fireSpread", "caribouMovement"),
  inputs = filelist,
  paths = list(modulePath = system.file("sampleModules", package = "SpaDES"))
)

mySim$landscape <- stack(mySim$DEM, mySim$forestAge, mySim$habitatQuality, mySim$percentPine)
```

### Inspect the new `simList` object

Use the simList accessor functions to determine the following:

1. the modules included in the simulation and their package dependencies;
2. the global parameter values used;
3. the module-specific parameter values used;
4. the simulation start and end times;
5. the timeunits used by the modules in the simulation;
6. the scheduled and completed event queues;
7. the objects (functions, data, etc.) used in the simulation;
8. the file paths used for simulation inputs and outputs.

```{r inspect-simList}
# list modules used in the simulation
modules(mySim)

# list module dependencies and packages used
depends(mySim)
packages(mySim)

# list global and module-specific param values
globals(mySim)
params(mySim)
P(mySim)       ## bonus: how do params() and P() differ?

# list start and end times
times(mySim)
start(mySim)
end(mySim)

# get the simulation and module timeunits
timeunit(mySim)
timeunits(mySim)

# get the scheduled and completed event queues
events(mySim)
completed(mySim)

# list the objects (functions, data, etc.) stored in the simList
objects(mySim)

# list the file paths used in the simulation
paths(mySim)
inputPath(mySim)
outputPath(mySim)
```

### Run the simulation

```{r run-sim}
dev()
mySimOut <- Copy(mySim) ## make a deep copy of the simList
mySimOut <- spades(mySimOut)
```

### Compare the simList before and after

```{r compare-sims}
times(mySim)
times(mySimOut)

objects(mySim)
objects(mySimOut)
```
