---
title: "SpaDES in Action: a forest succession, a wildfire models, and a caribou models"
author: "Tati Micheletti"
date: "Jan 18, 2019"
output:
  html_document: default
  pdf_document: default
---
```{r init, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, eval = FALSE)
```

# Integrated simulation models: forest succession, wildfire and caribou

## Brief description of the models used  

The collection of models is an exercise to understand how SpaDES works when integrating different modules. 
It is composed of 3 main models (with one to several submodules within each):  

1. [LandR](https://github.com/PredictiveEcology/Biomass_core/tree/Workshop): this model is composed of 3 main modules:  
    `Biomass_borealDataPrep`: This module converts open datasets that are available for all of Canada's forests, into the input requirements for LandR_BiomassCore, a forest landscape succession model derived from the Landis-II Biomass Succession Model. This has been partially tested for some parts of the Western Boreal Forest.
Specifically, it takes the Ecozone, Ecoprovince, and Ecoregion maps of Canada, and species specific biomass maps of Canada (from Beaudoin *et al.* 2014). Keeping these data preparations outside of the LandR_BiomassCore module maintains the modularity of LandR's Biomass_core. See more information on this model in [PredictiveEcology/Biomass_borealDataPrep](https://github.com/PredictiveEcology/Biomass_borealDataPrep)
    `Biomass_regeneration`: Biomass_regeneration is a SpaDES module that simulates post-disturbance regeneration mechanisms for Biomass_core.
As such, this module is mostly based on the post-disturbance regeneration mechanisms present in LANDIS-II Biomass Succession v3.2.1 extension (see [LANDIS-II Biomass Succession v3.2 User Guide](https://github.com/LANDIS-II-Foundation/Extension-Biomass-Succession/blob/master/docs/LANDIS-II%20Biomass%20Succession%20v3.2%20User%20Guide.docx) and [Scheller and Mladenoff (2004)](https://pdfs.semanticscholar.org/4d38/d0be6b292eccd444af399775d37a757d1967.pdf).
At the moment, the Biomass_regeneration module only simulates post-fire disturbance effects on forest species, by simulating post-fire mortality and activating serotiny or resprouting mechanisms for each species, depending on their traits (i.e. ability to resprout and/or germinate from seeds, serotiny, following fire).
Post-fire mortality behaves in a stand-replacing fashion, i.e. should a pixel be within a fire perimeter (determined by a fire raster) all cohorts see their biomasses set to 0.  
As for post-fire regeneration, the module first evaluates whether any species present prior to fire are serotinous.
If so, these species will germinate depending on light conditions and their shade tolerance, and depending on their (seed) establishment probability (i.e. germination success) in that pixel.
The module then evaluates if any species present before fire are capable of resprouting. If so the model growth these species depending, again, on light conditions and their shade tolerance, and on their resprouting probability (i.e. resprouting success).
For any given species in any given pixel, only serotiny or resprouting can occur.
Hence, species that are capable of both will only resprout if serotiny was not activated.  
In LANDIS-II, resprouting could never occur in a given pixel if serotiny was activated for one or more species.
According to the manual:

> If serotiny (only possible immediately following a fire) is triggered for one or more species, then neither resprouting nor seeding will occur.
> Serotiny is given precedence over resprouting as it typically has a higher threshold for success than resprouting.
> This slightly favors serotinous species when mixed with species able to resprout following a fire.
([LANDIS-II Biomass Succession v3.2 User Guide](https://github.com/LANDIS-II-Foundation/Extension-Biomass-Succession/blob/master/docs/LANDIS-II%20Biomass%20Succession%20v3.2%20User%20Guide.docx))

This is no longer the case in Biomass_regeneration, where both serotinity and resprouting can occur in the same pixel, although not for the same species. We feel that this is more realistic ecologically, as resprouters will typically regenerate faster  after a fire, often shading serotinous species and creating interesting successional feedbacks (e.g. light-loving serotinous species having to "wait" for canopy gaps to germinate). See more information on this model in [PredictiveEcology/Biomass_regeneration](https://github.com/PredictiveEcology/Biomass_regeneration)
    `Biomass_core`: A biomass succession model derived and modified from LANDIS-II Biomass Succession v3.2.1.
Biomass_core simulates forest succession in a spatially explicit manner (per pixel) by reproducing the population dynamics (growth and mortality), dispersal and competition for light resources of tree species. 
Like in LANDIS-II, dynamics are simulated in terms of their biomass per cohort (a combination of species and age), following a `biomass ~ age` curve  that is influenced by species-specific parameters such as growth, mortality, maximum biomass, and species traits such as longevity and tolerance to shade (see [LANDIS-II Biomass Succession v3.2 User Guide](https://github.com/LANDIS-II-Foundation/Extension-Biomass-Succession/blob/master/docs/LANDIS-II%20Biomass%20Succession%20v3.2%20User%20Guide.docx) and [Scheller and Mladenoff (2004)](https://pdfs.semanticscholar.org/4d38/d0be6b292eccd444af399775d37a757d1967.pdf).
One of three dispersal algorithms are available: 'no dispersal', 'universal dispersal'(every species can disperse across the entire landscape) and 'ward dispersal' (species disperse according to a dispersal kernel, with mean and maximum distances taken from LANDIS-II trait tables), the latter being used by default.
The biggest differences between Biomass_core and LANDIS-II lie on how the model is parametrised and initialized and how different processes have been compartmentalised in order for higher flexibility. See more information on this model in [PredictiveEcology/Biomass_core](https://github.com/PredictiveEcology/Biomass_core/tree/development)  

2. [scfm](https://github.com/tati-micheletti/scfm/tree/development) (Steve Cumming's fire model): this model is composed of 3 main modules:
  `ignition`  
  `escape`  
  `spread`  
  `scfm` is described by Cumming et al. (1998), with a more accessible version in Armstrong and Cumming (2003).
It was recently implemented as a collection of SpaDES modules by Cumming, McIntire, and Eddy (in prep), with the addition of automated parameter estimation for fire management records. Scfm models fire as a three stage stochastic process of ignition, escape and spread, with each process represented by a dedicated module. Ignition and spread have already been described. The escape stage models the effect of fire suppression, or other ecological or sampling effects that alter the lower end of the fire size distribution. The empirical quantity is the “escape probability”, the probability that a fire will attain a final size greater than the size of a single pixel. This is simulated in the model by determining an initial spread probability such that the probability that a fire stays within its cell of origin equals the escape probability, after accounting for the effect of lakes and other non-flammable geographic features on fire spread. In effect, this distinguishes the first step of the iterative fire spread process. These three modules are used to simulate a fire regime, in terms of the number of fires that start, the escape probability, and the mean fire size.    
In addition to the scfmIgnition, scfmEscape and scfmSpread modules, scfm makes use of several support modules (or objects passed by the user, as in this case) involved in parameter estimation and model calibration. Fire regime parameters for ignition, escape,  and a mean fire size, are estimated from historic fire data obtained from the Canadian National Fire Database (Canadian Forest Service, n.d.). Spread probabilities are tuned to replicate the empirical mean fire size by a newly implemented calibration procedure. First, the study region raster is buffered by a set distance and a flammability map is generated for the buffered region using the 2005 Land Cover Map of Canada (Latifovic et al., 2005). Landcover classes such as open water,  rock, and ice are classed as non-flammable.  Next, several thousand fires are ignited at random locations in landscape and spread with probabilities randomly sampled from a given range (typically 0.18 to 0.24) of spread probabilities. Fires do not start within the buffer area, but may spread to and from the buffer. This effectively removes the influence of edge effects in determining mean fire size, provided the buffer is “wide enough”.  Then the spread probabilities and resulting fire sizes are fit with a shape-constrained additive model (SCAM) (Pya and Wood, 2015). The SCAM is monotonic to ensure fire size increases for any incremental increase in spread probability. Lastly, a function minimizer, coupled with the `scam predict` method, is used to find the spread probability that will reproduce the estimated mean fire size for the region. Parameterization and calibration of the model can be done separately for all the polygons in a given shapefile (e.g., by ecoregion), allowing spatial variation in fire regime parameters among regions.  

3 .  [caribouRSF](https://github.com/tati-micheletti/caribouRSF/tree/master) (resource selection function): Statistical relationships between caribou telemetry locations and habitat characteristics have been summarized (Environment Canada 2011) by resource selection function (RSF; Manly et al. 2002) type models for various parts of Canada. We used the top RSF-type model, a binomial GLMM, for the Taiga Plains ecozone (Environment Canada 2011; Table 46). This model was estimated from individual caribou location data collected from 24 Alberta, 50 British Columbia, and 169 NWT VHF-collared adult female caribou between 2000 and 2010 (Environment Canada 2011; Table 39). The analysis used a habitat use-availability design, so the model predicts relative, not absolute, resource selection (Manly et al. 2002).  
In the Taiga Plains ecozone, relative resource selection by adult female boreal caribou is best described by:  
$$selection Probability \sim Elevation + Elevation^2 + Vrug + Vrug^2 +$$ 
$$RoadDensity + Deciduous + Shrub + Herb + Water + RecentBurn + OldBurn$$ [Eq. 1]  

The definitions and units of measurement for each covariate can be found in Table 40 of Environment Canada (2011). 
The methodology is described in Appendix 7.3 of Environment Canada (2011). 
The model [Eq. 1] includes both static and dynamic covariates. Static covariates were determined from initial conditions. Dynamic covariates were 
derived from raster layers maintained by one of the ecological process modules (modules listed above). Covariate types and sources are provided below:   
  1.  Elevation (static; AdaptWest)  
  2.  Vrug (static; topographic ruggedness (Sappington et al. 2007) taken from AdaptWest.  
  3.  RoadDensity (static; `anthropogenic`)  
  4.  Deciduous (dynamic; `LandR`)  
  5.  Shrub (static; LCC2005)  
  6.  Herb (static; LCC2005)  
  7.  Water (static; Ducks Unlimited Canada Hybrid Wetland Layer)  
  8.  RecentBurn (dynamic;  `scfm-fireSense`/`scfm`)  
  9.  OldBurn (dynamic;  `scfm-fireSense`/`scfm`)  
The RSF model is applied on a decadal time step. The output is a raster layer of relative resource selection across the study area. A map of model prediction error (standard error) is also produced. This error comes from bootstrapping the error in the model covariates.

## Running the simulations   

### 1. [Global script](vignettes/01_globalScript.html)

  We will present here minimal examples of how each one of these models work (i.e. `LandR`, `scfm`, and `caribouRSF`) and how to integrate these into one analysis.

### 2. [Installing necessary libraries and setting options](vignettes/02_librariesAndOptions.html)  

```{r librariesAndOptions}
library("usefun")
library("SpaDES")
library("LandR")
library("pemisc")
library("raster")
library("data.table")
options(
  "spades.recoveryMode" = 2,
  "spades.lowMemory" = TRUE,
  "LandR.assertions" = FALSE,
  "LandR.verbose" = 1,
  "reproducible.useMemoise" = FALSE, # Brings cached stuff to memory during the second run
  "reproducible.useNewDigestAlgorithm" = TRUE,  # use the new less strict hashing algo
  "reproducible.useCache" = TRUE,
  "pemisc.useParallel" = FALSE
)

```

### 3. [Project structure](vignettes/03_projectStructure.html)  

### 4. [Setting paths](vignettes/04_settingPaths.html)

```{r settingPaths}
inputDirectory <- checkPath(file.path(getwd(), "inputs"), create = TRUE)
outputDirectory <- checkPath(file.path(getwd(), "outputs"), create = TRUE)
modulesDirectory <- checkPath(file.path(getwd(), "modules"), create = TRUE)
cacheDirectory <- checkPath(file.path(getwd(), "cache"), create = TRUE)

setPaths(cachePath = cacheDirectory,
         modulePath = c(modulesDirectory, 
                        file.path(modulesDirectory, "scfm/modules")),
         inputPath = inputDirectory,
         outputPath = outputDirectory)
```

### 5. [Time](vignettes/05_time.html)

```{r time}
times <- list(start = 0, end = 10)
```

### 6. [Modules](vignettes/06_modules.html)

These are the modules we will use, in four different simulations.

```{r modules}

modulesLandR <- c("Biomass_borealDataPrep",
                  "Biomass_core")

modulesFire <- c("scfmLandcoverInit",
                 "scfmRegime",
                 "scfmDriver",
                 "scfmIgnition",
                 "scfmEscape",
                 "scfmSpread")

allModules <- c(modulesLandR,
                modulesFire,
                "caribouRSF")
```
 
 Note that we previously ran three auxiliary fire modules and saved the objects created. The parametrization of SCFM for the study area we have takes about 30 minutes.

### 7. [Parameters](vignettes/07_parameters.html)

```{r}
successionTimestep <- 1L
parameters <- list(
  Biomass_borealDataPrep = list(
    "subsetDataBiomassModel" = TRUE,
    "subsetDataAgeModel" = TRUE
  ),
  Biomass_core = list(
    ".plotInitialTime" = times$start
    , "sppEquivCol" = "Boreal"
    , "successionTimestep" = successionTimestep*10
    , ".useParallel" = FALSE
  ),
  Biomass_regeneration = list(
    "fireTimestep" = successionTimestep,
    "fireInitialTime" = times$start
  ),
  scfmLandcoverInit = list(
    ".plotInitialTime" = times$start
  ),
  scfmDriver = list(
    ".useParallel" = TRUE
  ),
    scfmRegime = list(
    "fireCause" = "L"
  ),
  scfmSpread = list(
    "pSpread" = 0.235,
    "returnInterval" = successionTimestep,
    "startTime" = times$start,
    ".plotInitialTime" = times$start+1,
    ".plotInterval" = successionTimestep
    ),
  caribouRSF = list(
    "decidousSp" = c("Betu_Pap", "Popu_Tre", "Popu_Bal"),
    "predictionInterval" = 20
  )
)
```

### 8. [Objects](vignettes/08_objects.html)  

#### 8a. Load and plot Study Area  

```{r studyArea}
# load studyArea
studyArea <- readRDS(file.path(getPaths()$inputPath, "studyArea.rds"))

# NWT land cover classes -- base map to locate study area and be the reference raster (i.e. resolution, crs and extent)
regionMap <- Cache(reproducible::prepInputs, 
                   url = "https://drive.google.com/open?id=1WhL-DxrByCbzAj8A7eRx3Y1FVujtGmtN",
                   targetFile = "LCC2005_V1_4a_BCR6_NWT.tif",
                   destinationPath = getPaths()$inputPath,
                   fun = "raster::raster", 
                   userTags = "objectName:regionMap")
Plot(regionMap, title = "The NWT"); Plot(studyArea, addTo = "regionMap", title = "The NWT")
vegetationMap <- Cache(reproducible::prepInputs, 
                          url = "https://drive.google.com/open?id=1WhL-DxrByCbzAj8A7eRx3Y1FVujtGmtN", 
                          studyArea = studyArea,
                          targetFile = "LCC2005_V1_4a_BCR6_NWT.tif",
                          destinationPath = getPaths()$inputPath,
                          fun = "raster::raster", 
                          userTags = "objectName:regionMapCropped", overwrite = TRUE)

# See LandR::sppEquivalencies_CA for current options for species names -- pick species names
#   in one column -- if not specified, it will pick all species in the studyArea, but will
#   try by downloading all the species in the boreal forest of Canada (lots of downloads, and likely 
#   wrong species selection)
sppNameVector = c("Pice_Gla", "Pice_Mar", "Pinu_Ban", "Lari_Lar", "Popu_Tre", "Betu_Pap")
    
```

#### 8b. Load objects from inputs folder [Optional]

```{r fireObjects}
flammableMap <- readRDS(file.path(getPaths()$inputPath, "flammableMap.rds"))
landscapeAttr <- readRDS(file.path(getPaths()$inputPath, "landscapeAttr.rds"))
scfmDriverPars <- readRDS(file.path(getPaths()$inputPath, "scfmDriverPars.rds"))
scfmRegimePars <- readRDS(file.path(getPaths()$inputPath, "scfmRegimePars.rds"))
firePoints <- readRDS(file.path(getPaths()$inputPath, "firePoints.rds"))
sppEquivalenciesTable <- readRDS(file.path(getPaths()$inputPath, "sppEquivalenciesTable.rds"))
```

#### 8c. Passing the study area to the `objects` R object  
```{r objects}
objects <- list(
"vegMap" = vegetationMap,
"sppNameVector" = sppNameVector,
"studyArea" = studyArea, 
"sppEquiv" = sppEquivalenciesTable, 
#  LCC05 = vegetationMap, # If you have no internet, you can supply these objects
#  "flammableMap" = flammableMap,
#  "landscapeAttr" = landscapeAttr,
#  "scfmDriverPars" = scfmDriverPars,
#  "scfmRegimePars" = scfmRegimePars,
#  "firePoints" = firePoints
"sppColorVect" = c(Betu_Pap = "#8DD3C7", Lari_Lar = "#F2F1BA", Pice_Gla = "#D6A2B0",
                     Pice_Mar = "#B19DAC", Pinu_Ban = "#E4B378", Popu_Tre = "#B3DE69",
                     Mixed = "#D0FB84")
  )
```

### 9. [SimInit and the simList](vignettes/09_simList.html)

We will run 3 different, independent, simulations: `LandR`, `scfm` and `caribou`, and at last, we will run a simulation with all the modules. The only differece in simulation calls is regarding modules (both `modules` argument and `loadOrder`) 

#### **FOREST SUCCESSION**

First, we will run the LandR (forest succession) model:  

```{r LandRrun}
clearPlot()
LandR <- simInitAndSpades(times = times,
                          objects = objects,
                          params = parameters,
                          modules = as.list(modulesLandR),
                          paths = getPaths(),
                          loadOrder = modulesLandR,
                          debug = 1)
```

#### **FIRE**

Then, we will run the fire model:

```{r scfmRun}
clearPlot()
scfm <- simInitAndSpades(times = times, 
                         objects = objects,
                         params = parameters,
                         modules = as.list(modulesFire),
                         paths = getPaths(),
                         loadOrder = modulesFire)
```

#### **CARIBOU RESOURCE SELECTION FUNCTION**

We will follow by running the caribou model:

```{r caribouRun}
caribou <- simInitAndSpades(times = times,
                            objects = objects,
                            params = parameters,
                            modules = as.list("caribouRSF"),
                            paths = getPaths(),
                            debug = 1)
```

The model threw the intended error. The caribou model was developed in a way to expect input data from either 1) the user, or 2) other modules. If this data is not provided, it stops and warns the user.

#### **INTEGRATED MODEL**

At last, we will integrate the forest succession, fire and caribou:

```{r integratedModel, message = FALSE, warning = FALSE}
clearPlot()
SpaDESinAction <- simInitAndSpades(times = times, 
                                   params = parameters, 
                                   modules = allModules, 
                                   objects = objects, 
                                   paths = getPaths(), 
                                   loadOrder = unlist(allModules), 
                                   debug = 1)
```
  
    
Note that you can see all events happening, as well as any messages printed during the simulation (especially when you use the argument `debug = 1`). Also really handy is to be able to see all plots while the simulation is running. This is helpful to visually identify potential problems even before the simulation ends. This is a great advantage when running SpaDES modules within RStudio. 

### 10. [Checking your outputs](vignettes/10_simOuts.html)

#### **FOREST SUCCESSION**

We can check some outputs from the forest succession module such as the `cohort table` (with information on species, age, biomass, etc.) and the map of `pixelGroups` (with  information on the specific spatial location of the cohort groups):  

```{r outputsLandR}
clearPlot()
plot(SpaDESinAction$pixelGroupMap)
knitr::kable(rbind(head(SpaDESinAction$cohortData), tail(SpaDESinAction$cohortData)))
```

#### **FIRE**

First, lets check the flammable map (where fire can burn), the last years' fires and the cummulative fire map:

```{r plotFire, include = FALSE}
# Fire plots
clearPlot()
Plot(SpaDESinAction$flammableMap, title = "Flammable Map", col = c("skyblue", "red"))
Plot(SpaDESinAction$rstCurrentBurn, title = "Last year's burns", col = c("lightgrey", "red"))
Plot(SpaDESinAction$burnMap, title = "Cummulative burn map")
```

We can also see some of the fire statistics such as:

1. Summary of burns
```{r fireSummary1}
# Fire Summary
knitr::kable(SpaDESinAction$burnSummary)
```

2. Probability of ignition of a pixel
```{r fireSummary2}
# Fire Summary
plot(SpaDESinAction$pIg)
```

3. Fire driver parameters for polygon 67, for example  

```{r fireSummary3}
# Fire Summary
SpaDESinAction$scfmDriverPars$`67`
```

#### **CARIBOU RESOURCE SELECTION FUNCTION**

We can also see some of the caribou outputs such as the original RSF coefficients' table, the caribou model, and predicted presence probability of caribou in years 0 and 10:

```{r caribouSummary1}
# Caribou Summary
knitr::kable(rbind(head(SpaDESinAction$caribouCoefTableRSF), tail(SpaDESinAction$caribouCoefTableRSF)))

cat(SpaDESinAction$caribouModelsRSF$TaigaPlains)

Plot(SpaDESinAction$predictedPresenceProbability$Year0$TaigaPlains$relativeSelection, title = "Predicted presence probability of caribou in Year 0")
Plot(SpaDESinAction$predictedPresenceProbability$Year10$TaigaPlains$relativeSelection, title = "Predicted presence probability of caribou in Year 10")
```

  
### Hands-on

Are you ready to start on SpaDES? 

## Happy SpaDESing!!!

