---
title: "Loading libraries and setting options"
author: "Tati Micheletti"
date: "21 January 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

#### A. Libraries  

The first step to run the simulations is to install all libraries options for it. For that, we can  use the function `Require` from the `reproducible` package. It installs packages that are not installed, and loads the installed package.

```{r libraries}
library("reproducible")
Require("usefun")
Require("SpaDES")
Require("LandR")
Require("pemisc")
Require("raster")
Require("data.table")
```

#### B. Options

The second step to run the simulations is to define some options for it:
`spades.recoveryMode`: One of the coolest latest features of SpaDES. Did your simulation stop in the middle because of a simple error. Avoid self-facepalming by setting how many events you want SpaDES to "save" before the current event and just restart from this point after fixing for errors (using `mySimulation <- restartSpades()`). No more waiting 3.5 hours for all the simulation to re-run because you forgot a `,`! 
`spades.lowMemory`: Useful when the computer you are working on is not loaded with RAM.
`LandR.assertions`: Turns on and off assertions. Might capture edge cases, but might also stop at false negatives.
`LandR.verbose`: Prints out all messages from the `LandR` package. Useful to understand what is happening.
`reproducible.useMemoise`:  Brings cached stuff to memory during the second run. Useful if you have enough memory RAM and not too much patience to wait for the simulation to re-run.
`reproducible.useNewDigestAlgorithm`: use the new less strict hashing algo.
`reproducible.useCache`: Do you want to wait over and over again for the same mechanistic process? Just cache it!
`pemisc.useParallel`: Do you have more than one core and enough memory? You can set this option to `TRUE` and take advantage of parallel processing.

```{r options}
options(
  "spades.recoveryMode" = 2,
  "spades.lowMemory" = TRUE,
  "LandR.assertions" = FALSE,
  "LandR.verbose" = 1,
  "reproducible.useMemoise" = TRUE,
  "reproducible.useNewDigestAlgorithm" = TRUE,
  "reproducible.useCache" = TRUE,
  "pemisc.useParallel" = FALSE
)
```

`SpaDES` has a lot more options that can be set. The sky is the limit to try them out!

