---
title: 'Setting paths: where everything starts'
author: "Tati Micheletti"
date: "21 January 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

It is time to set the paths for our folders, including the `modules` folder. Here we use the functions `checkPath()` from the package `reproducible` to create paths if these don't exist, and assign (`<-`) to an R object these paths.  

The function `setPaths()` from the `SpaDES.core` package sets the paths, which can be retrived at any time with the function `getPaths()` Note that we need to pass both folders that have modules (orginal module's folder AND the module's folder inside `scfm`'s directory)

```{r settingPaths}
inputDirectory <- checkPath(file.path(getwd(), "inputs"), create = TRUE)
outputDirectory <- checkPath(file.path(getwd(), "outputs"), create = TRUE)
modulesDirectory <- checkPath(file.path(getwd(), "modules"), create = TRUE)
cacheDirectory <- checkPath(file.path(getwd(), "cache"), create = TRUE)

setPaths(cachePath = cacheDirectory,
         modulePath = c(modulesDirectory, 
                        file.path(modulesDirectory, "scfm/modules")),
         inputPath = inputDirectory,
         outputPath = outputDirectory)

```
