---
title: 'simList: What is it and what is in it?'
author: "Alex M Chubaty, Eliot McIntire & Tati Micheletti"
date: "January 2019"
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(magrittr)
library(SpaDES)
```

A simulation in SpaDES happens in two phases: initializatoin (function `simInit()`, and the simulation call, function `spades()`). It is also possible to use both calls in one function: the `simInitAndSpades()`.  

The initialization function needs as inputs the 1) paths, 2) times, 3) parameters, 4) objects, and 5) modules, apart from the 6)  *loadOrder* of the modules. If the load order is not provided, the modules might be loaded in a different order than these are supposed to, which might return an error due to missing objects.

### Where does `simList` come from?

1. Initializing a simulation using `simInit()` creates a `simList` object.  
2. Running a simulation via a `spades()` call returns a modified `simList` object.  

### How are simulations specified?

A `simList` object is simply a structured data type containing various elements of a simulation.

The main components of a `simList` are:

1. A list of modules used;
2. The event queue;
3. A description of the data (object) dependencies.

### How `SpaDES` works: simLists

We can examine the `simList` object structure in the usual R fashion by printing (`show`ing) it, or by using `str()`:

```{r simlist-structure, echo=TRUE, eval=FALSE}
emptySim <- simInit()
emptySim  # same as show(emptySim)
str(emptySim)
```

**NOTE:** `simList`s are S4 objects, so we can use `getSlots()` and `slotNames()` to examine the object.

See also `?'.simList-class'`

### Accessing the parts of a simLists

<div class="columns-2">
```{r simList-accessors, echo=FALSE, eval=TRUE}
emptySim <- simInit()
slots <- paste0(slotNames(emptySim))
accessors <- c("modules", "params", "events", "current", "completed", "depends",
               "times", "inputs", "outputs", "paths", "envir", "objects") %>% 
  paste0("()")
data.frame(slot = slots, accessor = accessors) %>%
  knitr::kable(format = "html", padding = 0)
```

```{r simList-accessors2, echo=FALSE, eval=TRUE}
other_accessors <- c("packages", "globals", "start", "end", "timeunit", "timeunits",
                     "paths", "cachePath", "inputPath", "outputPath", "modulePath") %>%
  paste0("()")
data.frame(other_accessors = other_accessors) %>%
  knitr::kable(format = "html", padding = 0)
```
</div>


### Inspect the empty `simList` object

1. What are the names of each of the slots in the simList?
2. What do each of these slots contain? Hint: see `?'.simList-class'`.

```{r inspect-emptySim}
str(emptySim)
slotNames(emptySim)
getSlots('simList')
?'.simList-class'
```

Examples and an exercise can be found [here](https://spades-workshops.predictiveecology.org/articlesOct2019/04c-The_simList.html).

