---
title: 'Checking outputs: the answer might be 42!'
author: "Tati Micheletti"
date: "21 January 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

### Inspect the `simList` object

Use the simList accessor functions to determine the following:

1. the modules included in the simulation and their package dependencies;
2. the global parameter values used;
3. the module-specific parameter values used;
4. the simulation start and end times;
5. the timeunits used by the modules in the simulation;
6. the scheduled and completed event queues;
7. the objects (functions, data, etc.) used in the simulation;
8. the file paths used for simulation inputs and outputs.

```{r inspect-simList}
# list modules used in the simulation
modules(mySim)

# list module dependencies and packages used
depends(mySim)
packages(mySim)

# list global and module-specific param values
globals(mySim)
params(mySim)
P(mySim)       ## bonus: how do params() and P() differ?

# list start and end times
times(mySim)
start(mySim)
end(mySim)

# get the simulation and module timeunits
timeunit(mySim)
timeunits(mySim)

# get the scheduled and completed event queues
events(mySim)
completed(mySim)

# list the objects (functions, data, etc.) stored in the simList
objects(mySim)

# list the file paths used in the simulation
paths(mySim)
inputPath(mySim)
outputPath(mySim)
```

### Run the simulation

```{r run-sim}
dev()
mySimOut <- Copy(mySim) ## make a deep copy of the simList
mySimOut <- spades(mySimOut)
```

### Compare the simList before and after

```{r compare-sims}
times(mySim)
times(mySimOut)

objects(mySim)
objects(mySimOut)
```

