---
title       : "Exercises: Simulation Experiments and Replication"
author      : "Alex M Chubaty & Eliot McIntire"
date        : "December 8, 2016"
output: 
  html_document: 
    number_sections: yes
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, eval = FALSE)
```

Often we want to run a simulation with a variety of parameter values as inputs. This is sometimes called a simulation experiment. The `experiment` function gives us an easy interface to this.

We will use the `establishment` module that we created in exercise 4a.

# Download a module

```{r experiment}
library(SpaDES)
module.path <- file.path(dirname(tempdir()), "modules")
downloadModule("establishment", path = module.path)
setPaths(modulePath = module.path)
library(ggplot2)

modules <- list("establishment")

mySim <- simInit(modules = modules)

# make sure the plotting device is clear
clearPlot()
spades(mySim)
```

# Create an experiment

1. Find out which parameters are in this module. (Hint: where is the module? Can you open it easily from R?) 
2. Pick one of the parameters and create a range of values for it. 
3. Following the structure indicated in example 4 of `?experiment`, use `experiment` to run a `spades` call for each parameter set (note that `experiment` is just a wrapper around `spades`). 
4. Inspect the experiment structure using `attributes(mySim)`. 
5. Create a plot showing on the x axis your parameter values, and on the y-axis something like `sum(mySim$establish[])`, *i.e.*, the number of pixels that had an establishment event. 

# Advanced - Create a two-parameter experiment

1. Repeat as above, but vary two parameters.

[See a solution here](https://github.com/PredictiveEcology/SpaDES-modules/blob/master/modules/establishment/establishment.Rmd)
